---
title: "Getting Started with mrpkit"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Getting Started with mrpkit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(digits = 2)
```

```{r setup, message = FALSE}
library(mrpkit)
```

This vignette presents the typical workflow of `mrpkit`, an R package for implementing Multilevel Regression with Post-stratification (MRP).


# Introduction

The aim of `mrpkit` is to perform multilevel regression with post-stratification (MRP; Gelman & Little, 1997; Park, Gelman, & Bafumi, 2004). By using this package, you follow a specific workflow for MRP. Hence, it creates a more reproducible and accountable environment for conducting MRP.

Unlike many R packages, `mrpkit` uses [R6 objects](https://adv-r.hadley.nz/r6.html) (Chang, 2019). For a user the main implication is that instead of calling a function to perform a task, we call a method embedded in an object using `object$method`. Typical usage of this package will create four objects: `SurveyData`, `QuestionMap`, `SurveyMap`, and `SurveyFit`. There are several methods in each of these objects that will be explained later in this vignette.  


# Workflow

The MRP workflow implemented by `mrpkit` is:

1. Prepare the data.
2. Create `SurveyData` objects for the sample and post-stratification datasets.
3. Create a `SurveyMap` object to match the values between the sample and post-stratification (population) datasets for each question.
4. Map the questions between the sample and post-stratification datasets.
5. Fit the model by creating a `SurveyFit` object.
6. Predict the probability of the outcome using the post-stratification dataset.
7. Estimate the probability of the outcome at a level of interest, for example, state/province level.
8. Create a summary with common comparisons 
9. Visualize the probability of the outcome.

A more detailed example of this workflow follows.

## Prepare the data

In this vignette, we use simulated survey sample and post-stratification datasets. These datasets display the preference of voters to the Box Party (BP), a political party in a fake country, Shape World, along with demographic variables. The population is designed using `declare_population` from `DeclareDesign` (Blair, Cooper, Coppock, & Humphreys, 2019) to set the features in the population.

The variables in the simulated survey dataset are:

```{r shape-survey}
head(shape_survey)
```

And the summarized simulated population is:

```{r shape-popn}
head(approx_voters_popn)
```
 

## Create a `SurveyData` object

`SurveyData` objects are used to represent both the sample and the post-stratification datasets, along with their metadata. The sample metadata consists of the survey's variables, the questions that were asked, and the levels of the responses. It is also optionally possible to include the survey weight and its design formula in this object. 

The most important methods for `SurveyData` objects are `SurveyData$new` and `SurveyData$print`. The `SurveyData$new` method is used to transform a regular data frame into a `SurveyData` object, and `SurveyData$print` is used to display the created `SurveyData` object in the console.

In this example, we will show the steps to create `SurveyData` objects using the `SurveyData$new` method. This method takes the metadata mentioned previously for its arguments:

1. `data`: The data frame containing the sample or population data.
Note that this method will automatically use each of the factor, character, and binary variables to create questions and responses if you do not specify the list of questions and responses manually. 
2. `questions`: The question asked in the questionnaire for each of those columns.
3. `responses`: The possible responses for each question.
4. `weights`: This is an optional argument of the survey or post-stratification/population weight. If the weight is specified as a column in the datasets, then this argument should be specified as a string. In this case, it is "wt" for `shape_survey` and `approx_voters_popn`.

It is possible to have an already summarized post-stratification dataset. If this is the case, then the weight would be the size of each cell. Alternatively, if the entire individual-level population is given, then this argument should be omitted, and it would be automatically specified as 1. 

5. `design`: This is an optional argument of the design formula of the survey, specified using the `survey` (Lumley, 2020) package notation.

Here we create a `SurveyData` object for the sample dataset.

```{r survey-object}
box_pref <- SurveyData$new(
  data = shape_survey,
  questions = list(
    age = "Please identify your age group",
    gender = "Please select your gender",
    vote_for = "Which party did you vote for in the 2018 election?",
    highest_educ = "Please identify your completed highest education",
    state = "Which State do you live in?",
    y = "If today is election day, will you vote for the Box Party?"
  ),
  responses = list(
    age = levels(shape_survey$age),
    gender = levels(shape_survey$gender),
    vote_for = levels(shape_survey$vote_for),
    highest_educ = levels(shape_survey$highest_educ),
    state = levels(shape_survey$state),
    y = c("no", "yes")
  ),
  weights = "wt",
  design = list(ids =  ~ 1)
)

box_pref$print()
```

Now we create a `SurveyData` object for the population dataset. Usually, the post-stratification data is from a large survey, for example, the American Community Survey (ACS) or Demographic and Health Survey (DHS). 

```{r poststrat-object}
popn_obj <- SurveyData$new(
  data = approx_voters_popn,
  questions = c(
    age_group = "Which age group are you?",
    gender = "Which gender are you identified?",
    vote_pref = "Which party do you prefer to vote?",
    education = "What is the highest grade or level of school you have completed",
    state = "Please identify the state where you live in"
  ),
  responses = list(
    age_group = levels(approx_voters_popn$age_group),
    gender = levels(approx_voters_popn$gender),
    vote_pref = levels(approx_voters_popn$vote_pref),
    education = levels(approx_voters_popn$education),
    state = levels(approx_voters_popn$state)
  ),
  weights = "wt",
  design = list(ids =  ~ 1)
)

# print it
popn_obj$print() 
```

In addition, sometimes, we want to access the properties of the `SurveyData` object, for example, the weights, the design, the questions and the number of questions, or the responses. The `SurveyData` object has methods that allow us to do so: `SurveyData$weights()`, `SurveyData$design()`, `SurveyData$questions()`, `SurveyData$n_questions()`, `SurveyData$responses()`. 

## Map the values between survey and post-stratification data using `QuestionMap` objects

From the metadata above, we can see that the survey (`box_pref`) and population (`popn_obj`) objects have different column labels and response levels. Specifically, the column label for age in `box_pref` is `age`, whereas, in `popn_obj`, the column label is `age_group`. This variable also has different levels; `box_pref` has seven levels of age, whereas `popn_obj` has only four levels of age. 

The column names and levels need to be aligned for MRP and this is accomplished with the `QuestionMap$new` method.

This method takes three arguments:

1. `name`: The name of the underlying construct. This will be used in the modeling stage.
2. `col_names`: A character vector of the column label in the survey and post-stratification objects. The column name of the survey should always be the first element, followed by the column name of the post-stratification object. This order is not interchangeable.  
3. `values_map`: The list of the mapped values between the survey and post-stratification object. If there is a meaningful ordering over the values, they should be sorted over that order, either descending or ascending.

Here is an example of how this method works:

```{r questions}
# Create QuestionMap$object for the question related to age
q_age <- QuestionMap$new(
  name = "age",
  col_names = c("age", "age_group"),
  values_map = list(
    "18-25" = "18-35",
    "26-35" = "18-35",
    "36-45" = "36-55",
    "46-55" = "36-55",
    "56-65" = "56-65",
    "66-75" = "66+",
    "76-90" = "66+"
  )
)

# Create QuestionMap$object for the question related to party preference
q_party_pref <- QuestionMap$new(
  name = "party_pref",
  col_names = c("vote_for", "vote_pref"),
  values_map = list(
    "Box Party" = "BP",
    "BP" = "BP",
    "Circle Party" = "CP",
    "CP" = "CP"
  )
)

# Create QuestionMap$object for the question related to gender
q_gender <- QuestionMap$new(
  name = "gender",
  col_names = c("gender", "gender"),
  values_map = data.frame(
    "male" = "m",
    "female" = "f",
    "nonbinary" = "nb"
  )
)

# Create QuestionMap$object for the question related to education
q_educ <- QuestionMap$new(
  name = "highest_education",
  col_names = c("highest_educ", "education"),
  values_map = list(
    "no high school" = "no high school",
    "high school" = "high school",
    "some college" = "some college",
    "associates" = "some college",
    "4-year college" = "4-years college",
    "post-graduate" = "post-grad"
  )
)

# Create QuestionMap$object for the question related to state
q_state <- QuestionMap$new(
  name = "state",
  col_names = c("state", "state"),
  values_map = list(
    "State A" = "A",
    "State B" = "B",
    "State C" = "C",
    "State D" = "D",
    "State E" = "E"
  )
)
```


## Using the `SurveyMap` object

The `SurveyMap` object holds the mapping between a set of items in the survey and post-stratification datasets. It takes the `SurveyData` objects, which in this case are `bp_pref` and `popn_obj`, together with labels and values that have been matched and provided in the `QuestionMap` objects. The mapped object specifies the correspondences in the variables that will be used when fitting the model.

The `SurveyMap` object has many methods. Here we demonstrate the `new`, `add`,
`delete`, `replace`, `mapping`, `tabulate`, and `fit` methods. Use `?SurveyMap`
to see the other available methods.

1. `SurveyMap$new`

    This method is used to initialize a new `SurveyMap` object. This takes `SurveyData` and `QuestionMap` objects as its argument. You can include all of the questions here, or add it incrementally using `add` method that is explained in point 2. 

```{r map-object}
# Create a new SurveyMap object
ex_map <-
  SurveyMap$new(sample = box_pref, population = popn_obj, q_age)

# example of mapping all of the questions at once
# ex_map <- SurveyMap$new(sample = box_pref, population = popn_obj,
#                        q_age, q_educ, q_gender, q_party_pref, q_state)
```

2. `SurveyMap$add`

    To get rid of the warning above, we could use this method to add other `QuestionMap` objects to the map.
    
```{r add}
# Add questions incrementally
ex_map$add(q_educ)
ex_map$add(q_gender, q_party_pref, q_state)

print(ex_map)
```

3. `SurveyMap$delete`

    Now, we have all of the questions in the post-stratification object mapped. Sometimes, we do not want to include all of the variables in the post-stratification matrix. For example, in this case, we want to exclude `party_pref` from the matrix using `delete` method.

```{r delete}
# We can also use the label instead of the object name
ex_map$delete("party_pref")
print(ex_map)
```

4. `SurveyMap$replace`

    Suppose that you changed your mind and want to use another level of matching for a certain question. For example, in question regarding education, q4, you want to change the "associates" level to be equal to "4-years college". To do this, you can firstly create a new `QuestionMap` object corresponding to that new level mapping. Secondly, you can use the `replace` method to change the question object. This method takes two arguments, namely `old question` as the first argument and `new question` as the second argument.
    
```{r replace}
# Create a new QuestionMap corresponding to the new level matching
q_educ_new <- QuestionMap$new(
  name = "highest_educ",
  col_names = c("highest_educ", "education"),
  values_map = list(
    "no high school" = "no high school",
    "high school" = "high school",
    "some college" = "some college",
    "associates" = "4-years college",
    "4-year college" = "4-years college",
    "post-graduate" = "post-grad"
  )
)

# replace the old question with new question
ex_map$replace(q_educ, q_educ_new)
print(ex_map)
```
    
5. `SurveyMap$mapping`

    Once you are happy with your map object, you can prepare the mapped data for model fitting with the `mapping` method.

```{r mapping}
ex_map$mapping()
```


6. `SurveyMap$tabulate`

    The next step is to prepare the post-stratification table using the `tabulate` method. If you want to only use a certain variable in the post-stratification matrix, then you should put that variable as the argument. For instance, here, we only use `age` in the post-stratification matrix.

```{r tab-1}
ex_map$tabulate("age")
```

If you want to include all of the variables in the post-stratification matrix, then the method should not take any arguments.
    
    
```{r tab-2}
ex_map$tabulate()
```


7. `SurveyMap$fit`

    Finally, we can fit the model using `fit` method. Currently, `mrpkit` natively supports `rstanarm::stan_glmer` and `rstanarm::stan_glm` (Goodrich, Gabry, Ali, & Brilleman, 2020), `lme4::glmer` (Bates, Maechler, Bolker, & Walker, 2015) , `brms::brm` (Bürkner, 2018). However, you can also specify your own function. In this case, it should give a `formula` argument and a `data` argument that accepts a data frame (the requirement for a formula may be relaxed in the future). In this example, we fit three models using `fun=rstanarm::stan_glmer`, `fun=lme4::glmer`, and `fun=brms::brm` with `age`, `gender`, and `education` as the predictors.

```{r  fitting}
# Example of using rstanarm::stan_glmer
fit1 <- ex_map$fit(
  fun = rstanarm::stan_glmer,
  formula = y ~ (1 | age) + (1 | gender) + (1 | highest_educ) + (1 | state),
  family = "binomial",
  refresh = 0,
  cores = 2)

# Example of using lme4::glmer
fit2 <- ex_map$fit(
  fun = lme4::glmer,
  formula = y ~ (1 | age) + (1 | gender) + (1 | highest_educ) + (1 | state),
  family = "binomial")

```

```{r fit-brm, eval = FALSE}
# Example using brms::brm
fit3 <- ex_map$fit(
  fun = brms::brm,
  formula = y ~ (1 | age) + (1 | gender),
  family = "bernoulli",
  refresh = 100,
  cores = 2)
```

The resulting object is _not_ the fitted model object created by the modeling function, but rather a `SurveyFit` object that contains the fitted model and provides useful methods for working with it.

## Using the `SurveyFit` object

At this stage, we already have the fitted model which has stored as a `SurveyFit` object. Using this object we can generate the predicted probabilities for the post-stratification cells, aggregate the small area estimation as needed (e.g., state or province estimates), and visualize the aggregated estimates. 

### Get the probability of the outcome from each poststratification cells

After creating the `SurveyFit` object, we can generate the predicted probability of the outcome of each post-stratification cell using the `SurveyFit$population_predict` method. It returns a matrix with rows that correspond to the columns of post-stratification data and columns that correspond to the posterior samples.

Furthermore, if the model fitting is done with one of the natively supported functions (`rstanarm::stan_glmer`, `lme4::glmer`, `brms::brm`), then this method does not take any arguments. However, if you used a custom function in the model fitting, then this method takes multiple arguments, including a custom prediction function 
(see `?SurveyFit` for details). In this example, we will show how to use this method for the `fit1` object created using `rstanarm::stan_glmer`.

```{r predict}
# predict the probability of voting for the Box Party using the fit1 model
poststrat_est_fit1 <- fit1$population_predict()

dim(poststrat_est_fit1)
nrow(fit1$map()$poststrat_data())
```


### Aggregate the probability of the outcome to a higher level of estimate

The next step is to generate the population estimate or group estimate using the `SurveyFit$aggregate` method. This method takes two arguments, namely the post-stratification estimate data and the variable whose level to which the estimated value would be aggregated to. In this example, we want to aggregate the estimated value by the level of `age` and `state`. If the variable is not specified, then this method will generate a population estimate.

```{r aggregate}

# aggregate the predicted value by age
age_estimation <- fit1$aggregate(poststrat_est_fit1, by = "age")
str(age_estimation)

# aggregate the predicted value by state
state_estimation <- fit1$aggregate(poststrat_est_fit1, by = "state")
str(state_estimation)

# generate the population estimate
popn_estimation <- fit1$aggregate(poststrat_est_fit1)
str(popn_estimation)
```

The returned object is a data frame with number of rows equal to the number of
posterior draws times the number of levels of the `by` variable. Each of the
numbers in the `value` column represents the estimated probability of the
outcome. In this example, `age_estimation` contains the probability of voting
for the Box Party for each age group for each posterior draw. 

```{r show-aggr}
# get the mean and sd for each age group
library(dplyr)
age_estimation %>%
  group_by(age) %>%
  summarize(mean = mean(value), sd = sd(value))
```


### Summarise and compare

The output of `aggregate` is a posterior predictive distribution for the
population (or sub-populations) of interest. This can be used with the plot
function to visualize. However, we also might wish to make a table summary of
the posterior mean and standard deviation, along with comparisons with raw
estimates and weighted estimates. We can use the `summary` method to do this.
This method takes the aggregated data frame and produces a data frame of mean and
standard deviation estimates for each group (if in the aggregated frame) or
overall in the population.

```{r summarize}
fit1$summary(age_estimation)
fit1$summary(state_estimation)
fit1$summary(popn_estimation)
```
The standard deviation for the weighted estimate (`wtd`) is created by wrapping
the `survey` package using the original design specified, and the raw standard
deviation is $\sqrt\frac{pq}{n}$.\

### Visualize it

Once we have the aggregated estimates, we can visualize them using the `plot` method. This method takes the aggregated data frame and additional_stats option (the default is wtd and raw estimates) as its arguments. It generates a violin plot of the aggregated estimate by the level of a certain variable. When additional statistics are specified (options are `raw`, `wtd` and `mrp`; more than one can be specified as a vector), these are added as points with error bars representing the 95% confidence interval.  
In this example, we display the estimated plot for `age` and `state` and a density plot for the population estimates. 


```{r age-vis, fig.width= 6, fig.height=4}
plot_age <- fit1$plot(age_estimation) 
plot_age
```

The plot above shows the distribution of people who vote for the Box Party for each age group. The plot implies that people who are aged 36-55 are most likely to vote for the BP.

The next plot displays the probability of voting for the BP by the state without using additional statistics. We learn that the probability of the BP winning the election is high since most people in almost all the states (around 75%) vote for that party. State B is the state with the lowest probability of vote for the BP. The distribution still ranged more than 50%. Note that this violin plot is generated using `ggplot2` (Wickham, 2016). Hence, it could be modified the same way `ggplot2`'s plot is modified. For example, here, we can add the title and change the theme of the plot.


```{r state-vis, fig.width= 6, fig.height=4}
library(ggplot2)
plot_state <- fit1$plot(state_estimation, additional_stats = "none") +
  ggtitle("Probability of voting the BP by state") +
  theme_bw()
plot_state
```

Lastly, we show the plot of the population estimate, with lines representing the mean weighted and raw estimates. About 70% of people in the Shape World will be likely to vote for the Box Party. 

```{r,popn_estimate, fig.width= 6, fig.height=4}
fit1$plot(popn_estimation, additional_stats = c("raw","wtd"))
```

## References


Bates, D., Maechler, M., Bolker, B., & Walker, S. (2015). Fitting Linear Mixed-Effects Models Using lme4. *Journal of Statistical Software*, 67(1), 1-48. \doi:10.18637/jss.v067.i01.
  
Blair, G., Cooper, J., Coppock, A., & Humphreys, M. (2019). Declaring and Diagnosing Research Designs. *American Political Science Review* 113(3): 838-859. URL https://declaredesign.org/declare.pdf.

Bürkner, P.C. (2018). Advanced Bayesian Multilevel Modeling with the R Package brms. *The R Journal*, 10(1), 395-411. \doi:10.32614/RJ-2018-017.

Chang, W. (2019). R6: Encapsulated Classes with Reference Semantics. R package version 2.4.1. https://CRAN.R-project.org/package=R6.
  
Gelman, A., & Little, T. C. (1997). Poststratification into many categories using hierarchical logistic regression. *Survey Methodology*, 23(2), 127–135.

Goodrich, B., Gabry, J., Ali, I., & Brilleman S. (2020). rstanarm: Bayesian applied regression modeling via Stan. R package version 2.21.1 https://mc-stan.org/rstanarm/.

Lumley, T. (2020) "survey: analysis of complex survey samples". R package version 4.0.

Park, D. K., Gelman, A., & Bafumi, J. (2004). Bayesian multilevel estimation with poststratification: State-level estimates from national polls. *Political Analysis*, 12(4), 375–385.

R Core Team. (2020). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
  
Wickham, H., François, R., Henry, L., & Müller, K. (2020). dplyr: A Grammar of Data Manipulation. R package version 1.0.2. https://CRAN.R-project.org/package=dplyr
  
Wickham, H. (2016). *ggplot2: Elegant Graphics for Data Analysis*. Springer-Verlag New York



